Deterministic and statistical truncated KdV models for anomalous waves induced by abrupt depth change

### 2018-09-27 LatexNotes/Notes.tex
NM

We start with the variable-depth KdV equation given by Johnson in Eq.(3.149) on page 272 in dimensionless form 
\begin{equation}
2 H_X + \frac{3}{D^{7/4}} H H_\xi + \frac{1}{3} D^{1/2} H_{\xi \xi \xi} = 0
\end{equation}
All variables above are dimensionless, and by working backwards we can figure out how they relate to the original, dimensional variables. Below, the dimensional quantities are indicated with an overbar wherever there is the possibility of ambiguity
\begin{align}
& H = D^{1/4} \eta = \left( \frac{d}{h_0} \right)^{1/4} \frac{\bar{\eta}}{\eps h_0} \\
& X = \eps x = \frac{\eps^{3/2}}{h_0} \bar{x} \\
& \xi = \frac{\eps^{1/2}}{\sqrt{d h_0}} (\bar{x} - c \bar{t})
\end{align}
Here $d$ is the variable depth (which is dimensional) $c = \sqrt{g d}$ is the local wave speed, and $D = d/h_0$ is the variable depth ratio.

There are a few important things to remember when we reformulate this equation
\begin{enumerate}
\item We must retain Johnson's choice to keep the combination $D^{1/4} \eta$ together because that quantity matches across the abrupt depth change (the raw surface displacement does not match).
\item We want to normalize the standard deviation (or energy) of $\eta$ and therefore want to keep Johnson's choice to scale surface displacement on $a = \eps h_0$.
\item We have to scale the spatial variable (called $x$ or $\xi$) on a characteristic wavelength, which will be a different value on either side of the abrupt depth change. We have to do this so that the assumption of spatial periodicity will be reasonably satisfied.
\end{enumerate}
To achieve 3, we can be guided by the characteristic wavelength from shallow water theory
\begin{equation}
\lambda_c = {c}/{f_c} = \sqrt{g d}/f_c
\end{equation}
where $f_c$ is a characteristic forcing frequency and $c = \sqrt{g d}$ is the local wavespeed (in shallow-water theory). It is tempting to simply set $f_c$ to the peak forcing frequency $f_p$, which would result in the characteristic wavelength $\lambda_c$ being the peak wavelength $\lambda_p$. Perhaps a better choice, though, is to let $\lambda_c$ be an integer multiple of $\lambda_p$ to allow scales larger than the peak wavelength (which are certainly present in the experiments). Therefore, let us choose
\begin{equation}
\lambda_c = M \sqrt{gd} / f_p
\end{equation}
where $M$ is an integer (on the order of 10 or so). Notice that Johnson's formulation already incorporates the correct scaling of $\sqrt{d}$ to define $\xi$, which will make things convenient.

>>>>

We start with the variable-depth KdV equation given by Johnson in Eq.(3.149) on page 272 in dimensionless form 
\begin{equation}
2 H_X + \frac{3}{D^{7/4}} H H_\xi + \frac{1}{3} D^{1/2} H_{\xi \xi \xi} = 0
\end{equation}
All variables above are dimensionless, and by working backwards we can figure out how they relate to the original, dimensional variables. Below, the dimensional quantities are indicated with an overbar wherever there is the possibility of ambiguity
\begin{align}
& H = D^{1/4} \eta = \left( \frac{d}{h_0} \right)^{1/4} \frac{\bar{\eta}}{\eps h_0} \\
& X = \eps x = \frac{\eps^{3/2}}{h_0} \bar{x} \\
& \xi = \frac{\eps^{1/2}}{\sqrt{d h_0}} (\bar{x} - c \bar{t})
\end{align}
Here $d$ is the variable depth (which is dimensional) $c = \sqrt{g d}$ is the local wave speed, and $D = d/h_0$ is the depth ratio.

There are a few important things to remember when we reformulate this equation:
\begin{enumerate}
\item We must retain Johnson's choice to keep the combination $D^{1/4} \eta$ together because that quantity matches across the abrupt depth change (the raw surface displacement does not match).
\item We want to normalize the standard deviation (or energy) of $\eta$ and therefore want to keep Johnson's choice to scale surface displacement on $a = \eps h_0$.
\item We have to scale the spatial variable (called $x$ or $\xi$) on a characteristic wavelength, which will be a different value on either side of the abrupt depth change. We have to do this so that the assumption of spatial periodicity will be reasonably satisfied on both sides.
\end{enumerate}
To achieve 3, we can use the characteristic wavelength from shallow water theory as a guide,
\begin{equation}
\lambda_c = {c}/{f_c} = \sqrt{g d}/f_c
\end{equation}
where $f_c$ is a characteristic forcing frequency and $c = \sqrt{g d}$ is the local wavespeed (in shallow-water theory). It is tempting to simply set $f_c$ to the peak forcing frequency $f_p$, which would result in the characteristic wavelength $\lambda_c$ being the peak wavelength $\lambda_p$. Perhaps a better choice, though, is to let $\lambda_c$ be an integer multiple of $\lambda_p$ to allow scales larger than the peak wavelength (which are certainly present in the experiments). We therefore choose
\begin{equation}
\lambda_c = M \sqrt{gd} / f_p
\end{equation}
where $M$ is an integer (on the order of 10 or so). Notice that Johnson's formulation already incorporates the correct scaling of $\sqrt{d}$ to define $\xi$, which will make things convenient.
